---
title: "Phylogenetic eigenvector maps (PEM): Theory and practice"
author: Guillaume Guénard & Pierre Legendre
institute: Université de Montréal -- Département de sciences biologiques
date: 2024-09-15
format:
  revealjs:
    transition: none
    transition-speed: slow
    incremental: true
editor: visual
editor_options:
  markdown:
    wrap: 80
---

```{r, results='hide'}
library(knitr)
source("PEM_seminar-aux.R")
load(file="linear.rda")
```

# Part I: Theory

## Equations {transition="slide"}

$$
\bar{x} = \frac{\sum_{i=0}^N x_i}{N}
$$

$$
\mathrm{e}^{i\pi} - 1 = 0
$$

$$
\cos^2(x) + \sin^2(x) = 1
$$

$$
\mathbf{X} \mathbf{u}_i = \lambda_i \mathbf{u}_i
$$

$$
\mathbf{X} \mathbf{v}_i = \sigma_i \mathbf{u}_i
$$

## SVD {transition="slide"}

```{r, echo=TRUE}
mat <- matrix(c(1,3,1,6,7,4),3,2)
tmp <- svd(mat)
mat %*% tmp$v
tmp$u %*% diag(tmp$d)
```

## The influence matrix

- Rows represent the vertices

- Columns represent the edges

- Notation: $\mathbf{B} = [b_{i,j}]$ (binary)

## Rows represent the vertices

```{r, fig.height=3, fig.width=6}
plot1_graph(linear, "\"")
```

## Rows represent the vertices

```{r, fig.height=3, fig.width=6}
plot1_vertex(linear, 1L)
```

## Rows represent the vertices

```{r, fig.height=3, fig.width=6}
plot1_vertex(linear, 2L)
```

## Rows represent the vertices

```{r, fig.height=3, fig.width=6}
plot1_vertex(linear, 3L)
```

## Rows represent the vertices

```{r, fig.height=3, fig.width=6}
plot1_vertex(linear, 4L)
```

## Rows represent the vertices

```{r, fig.height=3, fig.width=6}
plot1_vertex(linear, 5L)
```

## Rows represent the vertices

```{r, fig.height=3, fig.width=6}
plot1_vertex(linear, 6L)
```

## Rows represent the vertices

```{r, fig.height=3, fig.width=6}
plot1_vertex(linear, 7L)
```

## Columns represent the edges

```{r, fig.height=3, fig.width=6}
plot1_graph(linear, "\"")
```

## Columns represent the edges

```{r, fig.height=3, fig.width=6}
plot1_edge(linear, 1L)
```

## Columns represent the edges

```{r, fig.height=3, fig.width=6}
plot1_edge(linear, 2L)
```

## Columns represent the edges

```{r, fig.height=3, fig.width=6}
plot1_edge(linear, 3L)
```

## Columns represent the edges

```{r, fig.height=3, fig.width=6}
plot1_edge(linear, 4L)
```

## Columns represent the edges

```{r, fig.height=3, fig.width=6}
plot1_edge(linear, 5L)
```

## Columns represent the edges

```{r, fig.height=3, fig.width=6}
plot1_edge(linear, 6L)
```

## The influence matrix

### Values

- The path between the root and vertex $i$ goes through edge $j$?
  
  > vertex $i$ is under the influence of edge $j$ $\implies$ $b_{i,j} = 1$
  
- Otherwise
  
  > $b_{i,j} = 0$

## Values of the influence matrix

```{r, fig.height=3, fig.width=6}
plot1_influence(linear, 1L, 1L)
```




# Part II: practice
